Basic Imaging: Viking Graben

### 2019-10-20 repro/papers/vikingbasic/paper.tex
added viking graben projects

\title{Basic Imaging of a 2D Marine Survey}
\maketitle
\label{ch:basic}
\inputdir{project}

@@

This section of the CAAM 641 course notes discusses so-called standard processing of the Viking Graben data, based on hyperbolic normal moveout (NMO). The connection between inversion and standard processing is the subject of the first part of this course. An earlier treatment can be found in my MSRI lecture notes from 2013, available on the TRIP web page, downloadable materials, short course and summer school materials, especially part 2.4. I also show how to transform the by-products of standard processing into suitable input for prestack imaging via RTM.

My workflow is recorded in the {\tt project/SConstruct} file. The reader should refer to this file for the detailed form of the commands used. The following sections outline the rationale for the choice of commands and their sequence. To reproduce my results, the reader will need to install SU, Madagascar, and TRIP packages. SU should be version 44 or later. The prestack part of the workflow requires parallel installation of TRIP, and at least 10 threads of execution, preferably more, to finish in reasonable time.

@@

The velocity-denominated quantity determined by flattening NMO-corrected CDP gathers is the so-called RMS velocity, or $v_{\rm rms}$, the root mean square of a local wave velocity expressed as a function of vertical travel time $t_0$:
\begin{equation}
\label{eqn:vrms}
v_{\rm rms}^2(t_0)= \frac{2}{t_0}\int_0^{t_0} d\tau \tilde{v}^2(\tau).
\end{equation}
Vertical traveltime is related to depth by 
\begin{equation}
\label{eqn:zt}
z=\int_0^{t_0}d\tau\tilde{v}(\tau)
\end{equation}
in which $\tilde{v}$ is the local (or {\em interval} wave velocity as a function of $t_0$. Velocity as a function of depth is then given by composing $\tilde{v}$ with the inverse of the tranformation (\ref{eqn:zt}).

@@

You will note the precence of many {\em diffraction tails} especially below 2 s. These may be partially collapsed either by poststack time migration using Gazdag's algorithm (Figure \ref{fig:paragazmig}) or, more effectively, by poststack depth migration using one of several methods. Gazdag poststack time migration assumes layered velocity structure, hence uses only the first column of the interval velocity as function of time. Nonetheless Figure \ref{fig:paragazmig} shows considerably less interference from diffraction tails; the graben structures that give this prospect its name are now clearly in evidence.

Poststack depth migration (a process that will be identified later as an approximation to the adjoint linearized zero-offset modeling operator) does even better - Figure \ref{fig:parapspi} displays the output of Phase-Shift-Plus-Interpolation poststack depth migration, using the velocity $v(z,x)$ depicted in Figure \ref{fig:paravintz} (obtained via use of the SU utility {\tt velconv}, see {\tt project/SConstruct} for details). The image gives reasonable depths for the various horizons. Also, this velocity model is the beginning of the next, prestack phase of Viking Graben processing.

With an eye to the prestack processing to be discussed later in the course, I include analogous results for downfiltered data. Jie Hou suggested a 5-10-30-40 Hz bandpass filter. The resulting AGC'd stack (Figure \ref{fig:parastackagcf}) shows the expected decrease in resolution over the original data, which has significant energy over 50 Hz.

@@

@@

The sequence of processing steps takes us far enough from the basic physics of wave propagation to raise questions about whether the images produced here are actually images of anything real. Later in the course, we turn to prestack processing based more directly on the acoustic wave equation - only a step further, but meaningful nonetheless. To telegraph part of the punch: prestack processing supports the validity of the images obtained in this paper by poststack processes. It seems that in the case of this ``tame'' data, classic seismic data processing does not steer you wrong.

\section{Suggested Projects}
\begin{enumerate}
\item The velocity analysis presented above seems quite minimal - only four analysis CMPs selected, more or less evenly spaced. Is there any advantage in choosing more analysis points? Does the stack become cleaner, the interval velocity any more believable?

\item Find another field data set and carry out a similar analysis.

\item Taking another tack, find a synthetic model with more lateral variation in velocity - Marmousi is an obvious choice - and carry out NMO based velocity analysis. Does the interval velocity so obtained bear any relation to the actual velocity (which of course you know, since the example is synthetic)? What about the PSPI migrated image and the actual model?

\item AGC is a crude tool for equalizing amplitudes, that completely destroys the relation between event and reflector dynamics. Is it possible to use a simpler scaling, such as by a power of $t$ or $z$, to  achieve the same enhancement of important reflectors at depth while maintaining a simple relation between reflector and reflection amplitudes?
\end{enumerate}

>>>>

\title{Basic Imaging: Viking Graben}
\date{}
\author{William W. Symes, The Rice Inversion Project}

\maketitle 
\parskip 12pt 
\begin{abstract}
Presents very simple processing sequence resulting in post-stack and prestack migrated images of Viking Graben data.
\end{abstract}

\inputdir{project}

@@

This paper discusses so-called standard processing of the Viking Graben data, based on hyperbolic normal moveout (NMO). The connection between inversion and standard processing is covered in detail in my MSRI lecture notes from 2013, available on the TRIP web page, downloadable materials, short course and summer school materials, especially part 2.4. I also show how to transform the by-products of standard processing into suitable input for prestack imaging via RTM.

My workflow is recorded in the {\tt project/SConstruct} file. The reader should refer to this file for the detailed form of the commands used. The following sections outline the rationale for the choice of commands and their sequence. To reproduce my results, the reader will need to install SU, Madagascar, and trip2.1 packages. SU should be version 44 or later. The prestack part of the workflow requires parallel installation of trip2.1, and at least 10 threads of execution, preferably more, to finish in reasonable time.

@@

As explained in the short course lectures, the velocity-denominated quantity determined by flattening NMO-corrected CDP gathers is the so-called RMS velocity, or $v_{\rm rms}$, the root mean square of a local wave velocity expressed as a function of vertical travel time $t_0$:
\begin{equation}
\label{eqn:vrms}
v_{\rm rms}^2(t_0)= \frac{2}{t_0}\int_0^{t_0} d\tau \tilde{v}^2(\tau).
\end{equation}
Vertical traveltime is related to depth by 
\begin{equation}
\label{eqn:zt}
z=\int_0^{t_0}d\tau\tilde{v}(\tau)
\end{equation}
in which $\tilde{v}$ is the local (or {\em interval} wave velocity as a function of $t_0$. Velocity as a function of depth is then given by composing $\tilde{v}$ with the inverse of the tranformation (\ref{eqn:zt}).

@@

You will note the precence of many {\em diffraction tails} especially below 2 s. These may be partially collapsed either by poststack time migration using Gazdag's algorithm (Figure \ref{fig:paragazmig}) or, more effectively, by poststack depth migration using one of several methods. Gazdag poststack time migration assumes layered velocity structure, hence uses only the first column of the interval velocity as function of tim. Nonetheless Figure \ref{fig:paragazmig} shows considerably less interference from diffraction tails; the graben structures that give this prospect its name are now clearly in evidence. Poststack depth migration (that is, what one would naturally mean - an approximation to the adjoint linearized zero-offset modeling operator) does even better - Figure \ref{fig:parapspi} displays the output of Phase-Shift-Plus-Interpolation poststack depth migration, using the velocity $v(z,x)$ depicted in Figure \ref{fig:paravintz} (obtained via use of the SU utility {\tt velconv}, see {\tt project/SConstruct} for details). The image gives reasonable depths for the various horizons. Also, this velocity model is the beginning of the next, prestack phase of Viking Graben processing.

With an eye to the prestack processing to be discussed in another paper, I include analogous results for downfiltered data. Jie Hou suggested a 5-10-30-40 Hz bandpass filter. The resulting AGC'd stack (Figure \ref{fig:parastackagcf}) shows the expected decrease in resolution over the original data, which has significant energy over 50 Hz.

@@

The sequence of processing steps takes us far enough from the basic physics of wave propagation to raise questions about whether the images produced here are actually images of anything real. In a subsequent paper, we turn to prestack processing based on the acoustic wave equation - only a step further, but meaningful nonetheless. To telegraph part of the punch: prestack processing supports the validity of the images obtained in this paper by poststack processes. It seems that in the case of this ``tame'' data, classic seismic data processing does not steer you wrong.

@@
